Categories of Optics

### 2018-03-12 paper.tex
Small updates

\usepackage{tikz-cd}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{cd}
\usetikzlibrary{arrows}

\usepackage{adjustbox}
\usepackage{etoolbox}
\BeforeBeginEnvironment {tikzpicture}{\begin{adjustbox}{valign=m}}
\AfterEndEnvironment {tikzpicture}{\end{adjustbox}}

@@

\newcommand{\C}{\mathscr{C}}
\newcommand{\homC}{\underline{\C}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\E}{\mathscr{E}}
\newcommand{\M}{\mathscr{M}}
\newcommand{\N}{\mathscr{N}}

@@

% Categories
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Cat}{\mathbf{Cat}}
\newcommand{\Prof}{\mathbf{Prof}}
\newcommand{\MonCat}{\mathbf{MonCat}}
\newcommand{\LaxMonCat}{\mathbf{LaxMonCat}}
\newcommand{\SymmMonCat}{\mathbf{SymmMonCat}}

@@

\begin{remark}
\todo{I hope I can get away with this...} In what follows, we use the standard cheat of omitting associativity isomorphisms
\end{remark}

Let $\C$ be a symmetric monoidal category, and suppose we have two pairs of objects $(S, S')$ and $(A, A')$. An \emph{optic} $p : (S, S') \hto (A, A')$ is an element of the set
\begin{align*}
\Optic_\C((S, S'), (A, A')) := \int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S')
\end{align*}
If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for this representative. Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.

\begin{remark}
Elements of $\Optic_\C((S, S'), (A, A'))$ have an appealing interpretation as string diagrams with a ``hole'' missing. We draw the pair $\langle l, r \rangle$ as
\begin{center}
\input{diagrams/generic-optic.tikz}
\end{center}
reading left to right, so the left half of the diagram represents $l$ and the right half $r$. The relation imposed by the coend takes the following form:
\begin{center}
\input{diagrams/coend-relation-left.tikz} \quad $=$ \quad \input{diagrams/coend-relation-right.tikz}
\end{center}
We will therefore omit the vertical cut point between $l$ and $r$ in subsequent diagrams; any choice yields a representative of same optic.
\end{remark}

@@

\begin{remark}
This functor is not necessarily faithful, see Remark \ref{lens-iota-not-faithful}
\end{remark}

@@

\begin{theorem}
$\Optic_\C$ is symmetric monoidal, where $(S, S') \otimes (T, T') = (S \otimes T, S' \otimes T')$, the unit is $(I, I)$, and the action on morphisms induced by
\begin{align*}
&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(T, N \otimes B) \times \C(N \otimes B', T')\\
\to \quad&\C(S \otimes T, M \otimes A \otimes N \otimes B) \times \C(M \otimes A' \otimes N \otimes B', S' \otimes T') && \text{(action of $\otimes$ on morphisms in $\C$)}\\
\to \quad&\C(S \otimes T, M \otimes N \otimes A \otimes B) \times \C(M \otimes N \otimes A' \otimes B', S' \otimes T') && \text{($\otimes$ in $\C$ is symmetric)}\\
\to \quad&\int^{M \in \C} \C(S \otimes T, M \otimes A \otimes B) \times \C(M \otimes A' \otimes B', S' \otimes T') && \text{(inclusion into coend)}
\end{align*}
\end{theorem}
\begin{proof}
Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:
\begin{align*}
\langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle
\end{align*}
This is functorial, as suggested by the equivalence of the following string diagrams:
\todo{todo}
%For future convenience, note that 

@@

%\langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
%&= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
%&= \langle l \otimes T, r \otimes T' \rangle

@@

%Now we check functoriality. Suppose we have:

@@

%\langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
%\langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
%\langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
%\langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')

@@

%with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}

@@

%&(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
%= \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%&\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
%&\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
%&\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%= \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
%= \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
%= \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
%= \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)

@@

@@

\begin{proposition}
Any isomorphism $p : (S, S') \hto (A, A')$ in $\Optic_\C$ is of the form $\iota(f, g)$ for two isomorphisms $S \to A$ and $A' \to S'$ in $\C$.
\end{proposition}
\begin{proof}

\end{proof}

There is some further useful structure in $\Optic_\C$. 
\begin{proposition}
\label{prop-costates}
The set of costates $(S, S') \hto (I, I)$ is isomorphic to $\C(S, S')$.
\end{proposition}
\begin{proof}
\begin{align*}
\Optic_\otimes((S, S'), (I, I)) 
&= \int^{M \in \C} \C(S, M \otimes I) \times \C(M \otimes I, S') \\
&\cong \int^{M \in \C} \C(S, M) \times \C(M, S') \\
&\cong \C(S, S')
\end{align*}
\end{proof}

@@

\begin{proposition}
Suppose the monoidal unit $I$ of $\C$ is terminal. Then the set of states $(I, I) \hto (A, A')$ is isomorphic to $\C(I, A)$.
\end{proposition}
\begin{proof}
First, note that
\begin{align*}
\Optic_\otimes((I,I), (A,A')) 
&= \int^{M \in \C} \C(I, M \otimes A) \times \C(M \otimes A', I) \\
&\cong \int^{M \in \C} \C(I, M \otimes A).
\end{align*}
Because the interior of this coend is mute in the contravariant position, the coend is equal to the colimit of the functor $\C(I, - \otimes A) : \C \to \Set$. But $\C$ has a terminal object, so $\int^{M \in \C} \C(I, M \otimes A) \cong \C(I, I \otimes A) \cong \C(I, A)$.
\end{proof}

@@

\begin{proposition}
A strong monoidal functor $F : \C \to \D$ induces a functor $\Optic(F) : \Optic_\C \to \Optic_\D$
\end{proposition}
\begin{proof}
On objects this is simply $\Optic(F)(S, S') = (FS, FS')$. The action on morphisms $\Optic_\C((S, S'), (A, A'))$ is induced by
\begin{align*}
&\C(S, M \otimes A) \times \C(M \otimes A', S) \\
\to &\D(FS, F(M \otimes A)) \times \D(F(M \otimes A'), FS) \\
\to &\D(FS, FM \otimes FA) \times \D(FM \otimes FA', FS) \\
\to &\int^{N \in \D} \D(FS, N \otimes FA) \times \D(N \otimes FA', FS).
\end{align*}
where we have used the strong monoidal isomorphism in both directions.
\todo{use equations instead}
\todo{argue for functoriality}
\end{proof}

@@

\begin{definition}[{\cite[Definition 5.1]{CoherenceForLenses}}]
A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
\begin{itemize}
\item A wide symmetric monoidal subcategory $\C_d$ of dualisable morphisms, with an involutive strong symmetric monoidal functor $(-)^* : \C_d^\op \to \C_d$
\item A symmetric monoidal extranatural family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$.
\end{itemize}
\end{definition}
\todo{The involution bit seems to be almost the same as asking for a symmetric monoidal dagger-category, that is stricter than this though}

>>>>

\usepackage{tikz-cd}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}
\usetikzlibrary{cd}
\usetikzlibrary{arrows}
\input{tikzcurverel}

@@

\newcommand{\C}{\mathscr{C}}
\newcommand{\homC}{\underline{\C}}
\newcommand{\D}{\mathscr{D}}
\newcommand{\E}{\mathscr{E}}
\newcommand{\M}{\mathscr{M}}
\newcommand{\N}{\mathscr{N}}
\newcommand{\T}{\mathscr{T}}

@@

% Categories
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Cat}{\mathbf{Cat}}
\newcommand{\Prof}{\mathbf{Prof}}
\newcommand{\MonCat}{\mathbf{MonCat}}
\newcommand{\LaxMonCat}{\mathbf{LaxMonCat}}
\newcommand{\SymmMonCat}{\mathbf{SymmMonCat}}
\newcommand{\Tele}{\mathbf{Tele}}

@@

\begin{remark}
\todo{I hope I can get away with this...} In what follows, we use the standard cheat of omitting associativity isomorphisms.
\end{remark}

\begin{definition}
Let $\C$ be a symmetric monoidal category, and suppose we have two pairs of objects $(S, S')$ and $(A, A')$. An \emph{optic} $p : (S, S') \hto (A, A')$ is an element of the set
\begin{align*}
\Optic_\C((S, S'), (A, A')) := \int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S')
\end{align*}
\end{definition}

If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for this representative. Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.

\begin{remark}
Elements of $\Optic_\C((S, S'), (A, A'))$ have an appealing interpretation as string diagrams with a ``hole'' missing. We draw the pair $\langle l, r \rangle$ as
\begin{center}
\input{diagrams/generic-optic.tikz}
\end{center}
reading left to right, so the left half of the diagram represents $l$ and the right half $r$. The relation imposed by the coend equates the following two diagrams:
\begin{center}
\input{diagrams/coend-relation-left.tikz} 
\qquad
\input{diagrams/coend-relation-right.tikz}
\end{center}
We will therefore omit the vertical cut point between $l$ and $r$ in subsequent diagrams; any choice yields a representative of same optic.
\end{remark}

@@

\begin{remark}
This functor is not necessarily faithful, see Remark \ref{lens-iota-not-faithful}.
\end{remark}

@@

%\begin{theorem}
%$\Optic_\C$ is symmetric monoidal, where $(S, S') \otimes (T, T') = (S \otimes T, S' \otimes T')$, the unit is $(I, I)$, and the action on morphisms induced by

@@

%&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(T, N \otimes B) \times \C(N \otimes B', T')\\
%\to \quad&\C(S \otimes T, M \otimes A \otimes N \otimes B) \times \C(M \otimes A' \otimes N \otimes B', S' \otimes T') && \text{(action of $\otimes$ on morphisms in $\C$)}\\
%\to \quad&\C(S \otimes T, M \otimes N \otimes A \otimes B) \times \C(M \otimes N \otimes A' \otimes B', S' \otimes T') && \text{($\otimes$ in $\C$ is symmetric)}\\
%\to \quad&\int^{M \in \C} \C(S \otimes T, M \otimes A \otimes B) \times \C(M \otimes A' \otimes B', S' \otimes T') && \text{(inclusion into coend)}

@@

%\end{theorem}
%\begin{proof}
%Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:

@@

%\langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle

@@

%This is functorial, as suggested by the equivalence of the following string diagrams:
%\todo{todo}
%%For future convenience, note that 
%%\begin{align*}
%%\langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
%%&= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
%%&= \langle l \otimes T, r \otimes T' \rangle
%%\end{align*}
%%Now we check functoriality. Suppose we have:
%%\begin{align*}
%%\langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
%%\langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
%%\langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
%%\langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')
%%\end{align*}
%%with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}
%%\begin{align*}
%%&(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
%%= \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%&\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
%%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
%%&\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
%%&\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
%%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%%= \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
%%= \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
%%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
%%= \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
%%= \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)
%%\end{align*}
%
%The structure morphisms are all lifted from the structure morphisms in $\C$:

@@

%\alpha_{(R, R'), (S, S'), (T, T')} &:= \iota(\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1}) \\
%\lambda_{(S, S')} &:= \iota(\lambda_{(S, S')}, \lambda_{(S, S')}^{-1}) \\
%\rho_{(S, S')} &:= (\rho_{(S, S')}, \rho_{(S, S')}^{-1}) \\
%s_{(S, S'), (T, T')} &:= \iota(s_{S, T}, s_{T', S'})

@@

% 
%Note that because $\iota(S, S') = (S, S')$, required equations for $\iota$ to be a monoidal functor hold by definition (although we don't yet know that $\Optic_\otimes$ is monoidal). The pentagon and triangle axioms then hold in $\Optic_\otimes$, as they are the image of the same axioms in $\C \times \C^\op$ under $\iota$. The only remaining thing to verify is that these structure maps are natural in $\Optic_\otimes$. 
%
%Let us check that $\alpha_{(R, R'), (S, S'), (T, T')}$ is natural in $(R, R')$. Suppose $f : (Q, Q') \hto (R, R')$ has representative $(l, r) : (Q \to M \otimes R, M \otimes R' \to Q')$, then:
%&\alpha_{(R, R'), (S, S'), (T, T')} \circ (f \otimes (S, S')) \otimes (T, T') \\
%&= \iota(\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1})  \circ (\langle l, r \rangle \otimes (S, S')) \otimes (T, T') \\
%&= \langle\lambda_{R \otimes (S \otimes T)}\alpha_{R,S,T}, \alpha_{R',S',T'}^{-1} \lambda^{-1}_{(R' \otimes S') \otimes T'} \rangle  \circ \left\langle (l \otimes S) \otimes T, (r \otimes S') \otimes T' \right\rangle \\
%&=\left\langle (M \otimes \lambda_{R \otimes (S \otimes T)}\alpha_{R,S,T})((l \otimes S) \otimes T), ((r \otimes S') \otimes T') (M \otimes \alpha_{R',S',T'}^{-1} \lambda^{-1}_{(R' \otimes S') \otimes T'} ) \right\rangle \\
%&=\left\langle (M \otimes \alpha_{R,S,T})((l \otimes S) \otimes T), ((r \otimes S') \otimes T') (M \otimes \alpha_{R',S',T'}^{-1} )\right\rangle \\
%&=\left\langle (l \otimes (S \otimes T))\alpha_{Q,S,T}, \alpha_{Q',S',T'}^{-1}(r \otimes (S' \otimes T')) \right\rangle \\
%&= \langle (I \otimes l \otimes (S \otimes T)) \lambda_{Q \otimes (S \otimes T)}\alpha_{Q,S,T}, (I \times r \otimes (S' \otimes T')) \alpha_{Q',S',T'}^{-1} \lambda^{-1}_{Q' \otimes (S' \otimes T')}\rangle \\
%&= \langle l \otimes (S \otimes T), r \otimes (S' \otimes T')\rangle \circ \left\langle \lambda_{Q \otimes (S \otimes T)}\alpha_{Q,S,T}, \alpha_{Q',S',T'}^{-1} \lambda^{-1}_{Q' \otimes (S' \otimes T')} \right\rangle \\
%&= \langle l, r \rangle \otimes ((S, S') \otimes (T, T')) \circ \iota(\alpha_{Q,S,T}, \alpha_{Q',S',T'}^{-1}) \\
%&= f \otimes ((S, S') \otimes (T, T')) \circ \alpha_{(Q, Q'), (S, S'), (T, T')} 
%\todo{There is some cheating going on here, composition actually adds an extra $\alpha$, but that wouldn't get in the way.}
%\end{proof}

\begin{theorem}
$\Optic_\C$ is symmetric monoidal, where $(S, S') \otimes (T, T') = (S \otimes T, T' \otimes S')$, the unit is $(I, I)$, and the action on morphisms is given by:
\begin{center}
\input{diagrams/tensor-on-morphisms.tikz}
\end{center}
\end{theorem}
\begin{proof}
Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:
\begin{align*}
\langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle
\end{align*}
This is functorial, as suggested by the equivalence of the following string diagrams:
\todo{todo}
%For future convenience, note that 
%\begin{align*}
%\langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
%&= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
%&= \langle l \otimes T, r \otimes T' \rangle
%\end{align*}
%Now we check functoriality. Suppose we have:
%\begin{align*}
%\langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
%\langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
%\langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
%\langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')
%\end{align*}
%with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}
%\begin{align*}
%&(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
%= \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%&\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
%&\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
%&\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
%= \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
%= \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
%&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
%= \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
%= \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)
%\end{align*}

@@

\begin{proposition}
$\iota : \C \times \C^\op \to \Optic_\C$ is a strong monoidal functor. \qed
\end{proposition}

%\begin{proposition}
%Any isomorphism $p : (S, S') \hto (A, A')$ in $\Optic_\C$ is of the form $\iota(f, g)$ for two isomorphisms $S \to A$ and $A' \to S'$ in $\C$.
%\end{proposition}
%\begin{proof}
%\todo{This might not actually be true..}
%\end{proof}

There is some further useful structure in $\Optic_\C$. 
\begin{proposition}
\label{prop-costates}
The set of costates $(S, S') \hto (I, I)$ is isomorphic to $\C(S, S')$.
\end{proposition}
\begin{proof}
\begin{align*}
\Optic_\C((S, S'), (I, I)) 
&= \int^{M \in \C} \C(S, M \otimes I) \times \C(M \otimes I, S') \\
&\cong \int^{M \in \C} \C(S, M) \times \C(M, S') \\
&\cong \C(S, S')
\end{align*}
\end{proof}

\todo{Is there a relationship between $\Optic_\C/(I, I)$ and $[2, \C]$?}

@@

\begin{proposition}
Suppose the monoidal unit $I$ of $\C$ is terminal. Then the set of states $(I, I) \hto (A, A')$ is isomorphic to $\C(I, A)$.
\end{proposition}
\begin{proof}
First, note that
\begin{align*}
\Optic_\C((I,I), (A,A')) 
&= \int^{M \in \C} \C(I, M \otimes A) \times \C(M \otimes A', I) \\
&\cong \int^{M \in \C} \C(I, M \otimes A).
\end{align*}
Because the interior of this coend is mute in the contravariant position, the coend is equal to the colimit of the functor $\C(I, - \otimes A) : \C \to \Set$. But $\C$ has a terminal object, so $\int^{M \in \C} \C(I, M \otimes A) \cong \C(I, I \otimes A) \cong \C(I, A)$.
\end{proof}

@@

\begin{proposition}
A monoidal functor $F : \C \to \D$ induces a  functor $\Optic(F) : \Optic_\C \to \Optic_\D$
\end{proposition}
\begin{proof}
On objects this is simply $\Optic(F)(S, S') = (FS, FS')$. On a morphism $\langle l, r \rangle : (S, S') \to (A, A')$, we define
\begin{align*}
F\langle l, r \rangle := \langle \phi^{-1}_{M,A} (Fl), (Fr) \phi_{M,A'}\rangle
\end{align*}
where $\phi_{M,A} : FM \otimes FA \to F(M \otimes A)$ is the structure map of the monoidal functor.
\todo{functoriality}
\todo{monoidality}
\end{proof}

@@

\begin{definition}[{\cite[Definition 5.1]{CoherenceForLenses}}]
A \emph{teleological category is} a symmetric monoidal category $(\C, \otimes, I)$, equipped with:
\begin{itemize}
\item A wide symmetric monoidal subcategory $\C_d$ of \emph{dualisable morphisms}, with an involutive strong symmetric monoidal functor $(-)^* : \C_d \to \overline{\C_d^\op}$, where---not finding a standard symbol for such a thing---we mean $\overline{\C_d^\op}$ to be the category with both the direction of the arrows \emph{and} the order of the tensor flipped: $(A \otimes B)^* = B^* \otimes A^*$.
\item A symmetric monoidal extranatural family of morphisms $\varepsilon_X : X \otimes X^* \to I$, called \emph{counits}, natural with respect to the morphisms in $\C_d$.
\end{itemize}
\end{definition}
Note that because $\C_d$ is symmetric monoidal and has the same collection of objects as $\C$, the symmetric monoidal structure of $\C$ must consist of dualisable maps.

Unpacking the definition, $\varepsilon$ being a symmetric monoidal extranatural transformation amounts to the following diagrams in $\C$ commuting:
\[
\begin{tikzcd}
X \otimes Y^* \ar[r, "f \otimes Y^*"]  \ar[d, "X \otimes f^*", swap] & Y \otimes Y^* \ar[d, "\varepsilon_Y"] \\
X \otimes X^* \ar[r, "\varepsilon_X", swap] & I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X^* \otimes X \ar[r, "\sigma"]  \ar[dr, "\varepsilon_{X^*}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd} \hspace{1cm}
\begin{tikzcd}
X \otimes Y \otimes Y^* \otimes X^* \ar[r, "X \otimes \varepsilon_Y \otimes X^*"]  \ar[dr, "\varepsilon_{X \otimes Y}", swap] & X \otimes X^* \ar[d, "\varepsilon_X"] \\
& I
\end{tikzcd}
\] 
where $f : X \to Y$ is in $\C_d$.

\begin{example}
Any compact closed category is a teleological category, where every morphism is dualisable and the unit morphisms have been forgotten.
\end{example}

\todo{\begin{example}
The funny graph example from the coherence paper?
\end{example}}

\begin{definition}
A \emph{teleological functor} $F : \C \to \D$ is a strong symmetric monoidal functor that restricts to a functor $F_d : \C_d \to \D_d$ on the dualisable subcategories, and such that $F(\varepsilon_X) = \varepsilon_{FX}$ \todo{up to the structure isomorphisms of $F$...}
\end{definition}
